### 2021-02-26 2021.csic5011/homework/homework02.tex
Add files via upload

\lecture{Homework 2. Random Matrix Theory and PCA}{Yuan Yao}{Due: 1.5 weeks}{Feb 26, 2020}

@@

\item *{\em Finite rank perturbations of random symmetric matrices:} Wigner's semi-circle law (proved by Eugene Wigner in 1951) concerns the limiting distribution of the eigenvalues of random symmetric matrices. It states, for example, that the limiting eigenvalue distribution of $n\times n$ symmetric matrices whose entries $w_{ij}$ on and above the diagonal $(i\leq j)$ are i.i.d Gaussians $\mathcal{N}(0,\frac{1}{4n})$ (and the entries below the diagonal are determined by symmetrization, i.e., $w_{ji}=w_{ij}$) is the semi-circle:
    $$p(t) = \frac{2}{\pi} \sqrt{1-t^2}, \quad -1\leq t \leq 1,$$
    where the distribution is supported in the interval $[-1,1]$.
\begin{enumerate}
\item Confirm Wigner's semi-circle law using MATLAB, Python, or R simulations (take, e.g., $n=400$).
\item Find the largest eigenvalue of a rank-1 perturbation of a Wigner matrix. That is, find the largest eigenvalue of the matrix $$W + \lambda_0 uu^T,$$ where $W$ is an $n\times n$ random symmetric matrix as above, and $u$ is some deterministic unit-norm vector. Determine the value of $\lambda_0$ for which a phase transition occurs. What is the correlation between the top eigenvector of $W+\lambda_0 uu^T$ and the vector $u$ as a function of $\lambda_0$? Use techniques similar to the ones we used in class for analyzing finite rank perturbations of sample covariance matrices.
\end{enumerate}

[Some Hints about homework] For Wigner Matrix\ $W=[w_{ij}]_{n\times n},w_{ij}=w_{ji},w_{ij}\sim N(0,\frac{\sigma}{\sqrt{n}})$, 
 the answer is
$$
\begin{array}{rcl}
\textrm{eigenvalue\ is}\ & &\lambda=R+\frac{1}{R}\\
\textrm{eigenvector\ satisfies}\ && (u^{T}\hat{v})^{2}=1-\frac{1}{R^{2}} \\
\end{array}
$$

>>>>

\lecture{Homework 2. Random Matrix Theory and PCA}{Yuan Yao}{Due: 1 week}{Feb 19, 2021}

@@

\item {\em Finite rank perturbations of random symmetric matrices:} Wigner's semi-circle law (proved by Eugene Wigner in 1951) concerns the limiting distribution of the eigenvalues of random symmetric matrices. It states, for example, that the limiting eigenvalue distribution of $n\times n$ symmetric matrices whose entries $w_{ij}$ on and above the diagonal $(i\leq j)$ are i.i.d Gaussians $\mathcal{N}(0,\frac{1}{4n})$ (and the entries below the diagonal are determined by symmetrization, i.e., $w_{ji}=w_{ij}$) is the semi-circle:
    $$p(t) = \frac{2}{\pi} \sqrt{1-t^2}, \quad -1\leq t \leq 1,$$
    where the distribution is supported in the interval $[-1,1]$.
\begin{enumerate}
\item Confirm Wigner's semi-circle law using MATLAB, Python, or R simulations (take, e.g., $n=400$).
\item Find the largest eigenvalue of a rank-1 perturbation of a Wigner matrix. That is, find the largest eigenvalue of the matrix $$W + \lambda_0 uu^T,$$ where $W$ is an $n\times n$ random symmetric matrix as above, and $u$ is some deterministic unit-norm vector. Determine the value of $\lambda_0$ for which a phase transition occurs. What is the correlation between the top eigenvector of $W+\lambda_0 uu^T$ and the vector $u$ as a function of $\lambda_0$? Use techniques similar to the ones we used in class for analyzing finite rank perturbations of sample covariance matrices.
\end{enumerate}

[Some Hints about homework] For Wigner Matrix\ $W=[w_{ij}]_{n\times n},w_{ij}=w_{ji},w_{ij}\sim {\mathcal N}(0,\frac{\sigma}{\sqrt{n}})$, 
 the answer is
$$
\begin{array}{rcl}
\textrm{eigenvalue\ is}\ & &\lambda=R+\frac{1}{R}\\
\textrm{eigenvector\ satisfies}\ && (u^{T}\hat{v})^{2}=1-\frac{1}{R^{2}} \\
\end{array}
$$
